Comparison of dimensionality reduction schemes for parallel global optimization algorithms \thanks{The study was supported by the Russian Science Foundation, project No 16-11- 10150}

### 2018-04-14 2018/RSD/MultiEvolvents/RSD.tex
Some style corrections

\begin{document}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
10150}}
%
\titlerunning{Dimension reduction schemes in parallel GO}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.

@@

In the present paper, the parallel algorithms for solving the multiextremal optimization problems
are considered. In the multiextremal problems, the opportunity of reliable estimate of the global
optimum is based principally on the availability of some information on the function known
{\textit a priori} allowing relating the probable values of the optimized function to the known
values at the points of performed trials. Very often, such an information on the problem being
solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
Lipschitz condition with the constant $L$ not known {\textit a priori} (see, for example,
\cite{Jones,Gablonsky,Evtushenko}). At that, the objective function could be defined by a
program code i. e. could represent a black-box-function. Such problems are presented in the
applications widely (problems of optimal design of objects and technological processes in
various fields of technology, problems of model fitting according to observed data in scientific
research, etc.).

Many methods destined to solving the problems of the class specified above reduce the solving
of a multidimensional problem to solving the one-dimensional subproblems implicitly (see, for
example, the methods of diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} or
simplicial partitions \cite{Zilinskas2008,Zilinskas2014}). In the present work, we will use the
approach developed in Lobachevskii State University of Nizhny Novgorod based on the idea of
the dimensionality reduction with the use of Peano space-filling curves $y(x)$ mapping the
interval $[0,1]$ of the real axis onto an $n$-dimensional cube continuously and unambiguously.

Several methods of constructing the evolvents approximating the theoretical Peano curve have
been proposed In Refs. \cite{strongin1978,Strongin1992,Gergel2009,Goryachih2017}. These
methods were implemented in the Globalizer software system \cite{globalizerSystem}. The goal
of the present study was comparing the properties of the evolvents and the selecting the most
suitable ones for the use in the parallel global optimization algorithms.

%\end{Russian}
%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
\(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
space.
\par
Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is time-consuming.
%As a result, the overall time of solving the optimization
%problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
%It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
%is highly important, since an estimate of the global minimum can be constructed on the
%basis of a finite number of computed values of the optimized function only in this case .

@@

Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978},
\cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the
following:
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by
using the mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point
\(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-
dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

@@

One of the possible ways to overcome the negative effects of using a numerical
approximation of evolvent (it destroys the information about the neighbour points in $\mathbb{R}^N$ space,
see \cite{Strongin1992}) consists in using the multiple mappings
\begin{equation}%\label{eq:142}
Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
\end{equation}
instead of single Peano curve $y(x)$ (see \cite{Strongin1991,Strongin1992,strSergGO}).

@@

As it has been already mentioned above (Sec.~\ref{sec:shifted}), the loss of information on the
proximity of the points in the multidimensional space could be compensated in part by the use
of multiple mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$. However, the Peano-type curve preserves
a part of this information itself: it is not an injective mapping. Therefore, if a single image
$y(x)\in \mathbb{R}^N$ is available, on can obtain several different $x$ preimages
$t_j\in[0,1], t_j \not = x$, which could be added into the search information of the index
method later.

The Peano-type curve used in (\ref{eq:oneDimTask}) for the dimensionality reduction is
defined via the transition to the limit. Therefore, it cannot be computed directly. In the
numerical optimization, some approximation of this curve is used, and it is an injective piecewise-
linear curve. In Ref.~\cite{strongin1978}, a non-univalent mapping of a uniform grid in the
interval $[0,1]$ onto a uniform grid in a hypercube $D$ has been proposed. Each
multidimensional node can have up to $2^N$ one-dimensional preimages. In
Fig.~\ref{fig:noninjective}, the grid in the $\mathbb{R}^2$ space is marked by the crosses, for
two nodes of which the corresponding one-dimensional preimages from $[0,1]$ are pointed
(marked by the squares and circles). Each node mentioned above has 3 preimages.

@@

\section{Parallel Computations for Solving Global Optimization Problems.}
\subsection{Core Multidimensional Algorithm of Global Search}

@@

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the
sequential algorithm (except the method of the trial execution). Each processor has its own copy
of the software realizing the computations of the problem functions and the decision rule of the
index algorithm. For the organization of the interactions among the processors, the queues are
created on each processor, where the processors store the information on the executed iterations
in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
\par
The proposed parallelization scheme was implemented with the use of MPI technology. Main
features of implementation consist in the following. A separate MPI-process is created for each
of \(S\) one-dimensional problems being solved, usually, one process per one processor
employed. Each process can use p threads, usually one thread per an accessible core.
\par
At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p
trials in parallel at the points \(x^{(s+i_S)},0\leqslant i<p\). At that, each process stores all
\(S_p\) points, and an attribute indicating whether this point is blocked by another process or
not is stored for each point. Let us remind that the point is blocked if the process starts the
execution of a trial at this point.
\par
At every iteration of the algorithm, operating within the \(s\)-th process, determines the
coordinates of \(p\) ``its own" trial points. Then, the interchange of the coordinates of images of
the trial points \(y^{(s+i_S)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to
each one). After that, the preimages \(x^{(q+i_S)},0\leqslant q<S,q\not=s\) of the points
received by the \(s\)-th process from the neighbor ones are determined with the use of the
evolvent \(y^s (x)\). The points blocked within the \(s\)-th process will correspond to the
preimages obtained. Then, each process performs the trials at the non-blocked points, the
computations are performed in parallel using OpenMP. The results of the executed trials (the
index of the point, the computed values of the problem functions, and the attribute of
unblocking of this point) are transferred to all rest processes. All the points are added to the
search information database, and the transition to the next iteration is performed.

@@

The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhny Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16
cores were available per a node). All considered algorithms and evolvents were implemented
using C++ within the Globalizer software system~\cite{globalizerSystem}.
In order to enable the parallelism,
OpenMP technology was used on a single node, and MPI standard --- for the parallelization on
several nodes.

@@

\bibitem{sergeyevStronginLera2013}% (MR3113120) [10.1007/978-1-4614-8042-6]
\newblock Y. D. Sergeyev, R. G. Strongin and D. Lera,
\newblock \emph{Introduction to Global Optimization Exploiting Space-filling Curves},
\newblock Springer, 2013.

\bibitem{strongin1978}% (MR509033)
\newblock R. G. Strongin,
\newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
\newblock Moscow: Nauka, In Russian, 1978.

\bibitem{Strongin1992}% (MR1263606) [10.1007/BF00122428]
\newblock R. G. Strongin,
\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems
employing a set of joint space-filling curves}},
\newblock \emph{J. Glob. Optim.}, \textbf{2} (1992), 357--378.

\bibitem{stronginGergelBarkalovParGO}
\newblock R. G. Strongin, V. P. Gergel, V. A. Grishagin and K. A. Barkalov,
\newblock \emph{Parallel Computations for Global Optimization Problems},
\newblock Moscow State University (In Russian), Moscow, 2013.

\bibitem{strSergGO}%5 (MR1797058) [10.1007/978-1-4615-4677-1]
\newblock R. G. Strongin and Y. D. Sergeyev,
\newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel
Algorithms},
\newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014).

\bibitem{zilinskTornGO}% (MR1100586) [10.1007/3-540-50871-6]
\newblock A. T\"orn and A. \v Zilinskas,
\newblock \emph{Global Optimization},
\newblock Springer, 1989.

\bibitem{Goryachih2017}
\newblock Goryachih, A.
\newblock \emph{A class of smooth modification of space-filling curves for global optimization
problems}
\newblock Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 57-65, 2017

\bibitem{zhigljavskyRandGO}% (MR1187048) [10.1007/978-94-011-3436-1]
\newblock A. A. Zhigljavsky,
\newblock \emph{Theory of Global Random Search},
\newblock Kluwer Academic Publishers, Dordrecht, 1991.

@@

\bibitem{globalizerSystem}
Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: \emph{A novel supercomputer software
system for solving time-consuming global optimization problems.} Numerical Algebra, Control
\& Optimization \textbf{8(1)}, 47--62, 2018

@@

\bibitem{Jones}
Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M.
(eds.) \emph{The Encyclopedia of Optimization}, 2nd edn., pp. 725--735. Springer, Heidelberg
(2009).

>>>>

\begin{document}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms 
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
10150}}
%
\titlerunning{Dimension reduction schemes in parallel GO}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhny Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.

@@

In the present paper, the parallel algorithms for solving the multiextremal optimization problems
are considered. In the multiextremal problems, the opportunity of reliable estimate of the global
optimum is based principally on the availability of some information on the function known
{\textit a priori} allowing relating the probable values of the optimized function to the known
values at the points of performed trials. Very often, such an information on the problem being
solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
Lipschitz condition with the constant $L$ not known a priori (see, for example,
\cite{Jones,Gablonsky,Evtushenko}). At that, the objective function could be defined by a
program code i. e. could represent a ``black-box''-function. Such problems are presented in the
applications widely (problems of optimal design of objects and technological processes in
various fields of technology, problems of model fitting according to observed data in scientific
research, etc.).

Many methods destined to solving the problems of the class specified above reduce the solving
of a multidimensional problem to solving the one-dimensional subproblems implicitly (see, for
example, the methods of diagonal partitions \cite{Sergeyev2006,SergeyevKvasov2015} or
simplicial partitions \cite{Zilinskas2008,Zilinskas2014}). In the present work, we will use the
approach developed in Lobachevsky State University of Nizhni Novgorod based on the idea of
the dimensionality reduction with the use of Peano space-filling curves $y(x)$ mapping the
interval $[0,1]$ of the real axis onto an $n$-dimensional cube continuously and unambiguously.

Several methods of constructing the evolvents approximating the theoretical Peano curve have
been proposed in \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. These
methods were implemented in the Globalizer software system \cite{globalizerSystem}. The goal
of the present study was comparing the properties of the evolvents and the selecting the most
suitable ones for the use in the parallel global optimization algorithms.

%\end{Russian}
%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
\(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
space.
\par
Usually, the objective function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is time-consuming.
%As a result, the overall time of solving the optimization
%problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
%It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
%is highly important, since an estimate of the global minimum can be constructed on the
%basis of a finite number of computed values of the optimized function only in this case .

@@

Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978},
\cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the % minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem
(\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the
following:
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by
using the mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point
\(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

@@

One of the possible ways to overcome the negative effects of using a numerical
approximation of evolvent (it destroys the information about the neighbour points in $\mathbb{R}^N$ space,
see \cite{Strongin1992}) consists in using the multiple mappings
\begin{equation}%\label{eq:142}
Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
\end{equation}
instead of single Peano curve $y(x)$ (see \cite{Strongin1992,strSergGO,Strongin1991}).

@@

As it has been already mentioned above (Sec.~\ref{sec:shifted}), the loss of information on the
proximity of the points in the multidimensional space could be compensated in part by the use
of multiple mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$. However, the Peano-type curve preserves
a part of this information itself: it is not an injective mapping. Therefore, if a single image
$y(x)\in \mathbb{R}^N$ is available, on can obtain several different preimages
$t_j\in[0,1], t_j \not = x$, which could be added into the search information of the method later.

The Peano-type curve used in (\ref{eq:oneDimTask}) for the dimensionality reduction is
defined via the transition to the limit. Therefore, it cannot be computed directly. In the
numerical optimization, some approximation of this curve is used, and it is an injective piecewise-linear curve. In \cite{strongin1978} a non-univalent mapping of a uniform grid in the
interval $[0,1]$ onto a uniform grid in a hypercube $D$ has been proposed. Each
multidimensional node can have up to $2^N$ one-dimensional preimages. In
Fig.~\ref{fig:noninjective}, the grid in the $\mathbb{R}^2$ space is marked by the crosses, for
two nodes of which the corresponding one-dimensional preimages from $[0,1]$ are pointed
(marked by the squares and circles). Each node mentioned above has 3 preimages.

@@

\section{Parallel Computations for Solving Global Optimization Problems.}
\subsection{Core Multidimensional Algorithm of Global Search (MAGS)}

@@

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the
sequential algorithm (except the method of the trial execution). Each processor has its own copy
of the software realizing the computations of the problem functions and the decision rule of the
index algorithm. For the organization of the interactions among the processors, the queues are
created on each processor, where the processors store the information on the executed iterations
in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
\par
The proposed parallelization scheme was implemented with the use of MPI technology. Main
features of implementation consist in the following. A separate MPI-process is created for each
of \(S\) one-dimensional problems being solved, usually, one process per one processor
employed. Each process can use $p$ threads, usually one thread per an accessible core.
\par
At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p
trials in parallel at the points \(x^{(s+iS)},0\leqslant i<p\). At that, each process stores all
\(S_p\) points, and an attribute indicating whether this point is blocked by another process or
not is stored for each point. Let us remind that the point is blocked if the process starts the
execution of a trial at this point.
\par
At every iteration of the algorithm, operating within the \(s\)-th process, determines the
coordinates of \(p\) ``its own" trial points. Then, the interchange of the coordinates of images of
the trial points \(y^{(s+iS)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to
each one). After that, the preimages \(x^{(q+iS)},0\leqslant q<S,q\not=s\) of the points
received by the \(s\)-th process from the neighbor ones are determined with the use of the
evolvent \(y^s (x)\). The points blocked within the \(s\)-th process will correspond to the
preimages obtained. Then, each process performs the trials at the non-blocked points, the
computations are performed in parallel using OpenMP. The results of the executed trials (the
index of the point, the computed values of the problem functions, and the attribute of
unblocking of this point) are transferred to all rest processes. All the points are added to the
search information database, and the transition to the next iteration is performed.

@@

The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhny Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16
cores were available per a node). All considered algorithms and evolvents were implemented
using C++ within the Globalizer software system~\cite{globalizerSystem}.
In order to enable the parallelism,
OpenMP was used on a single node, and MPI was used for the parallelization on
several nodes.

@@

\bibitem{sergeyevStronginLera2013}% (MR3113120) [10.1007/978-1-4614-8042-6]
\newblock Y. D. Sergeyev, R. G. Strongin and D. Lera,
\newblock \emph{Introduction to Global Optimization Exploiting Space-filling Curves},
\newblock Springer (2013)

\bibitem{strongin1978}% (MR509033)
\newblock R. G. Strongin,
\newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
\newblock Moscow: Nauka (1978) (In Russian)

\bibitem{Strongin1992}% (MR1263606) [10.1007/BF00122428]
\newblock R. G. Strongin,
\newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems
employing a set of joint space-filling curves}},
\newblock \emph{J. Glob. Optim.}, \textbf{2}, 357--378 (1992)

\bibitem{stronginGergelBarkalovParGO}
\newblock R. G. Strongin, V. P. Gergel, V. A. Grishagin and K. A. Barkalov,
\newblock \emph{Parallel Computations for Global Optimization Problems},
\newblock Moscow State University, Moscow (2013) (In Russian)

\bibitem{strSergGO}%5 (MR1797058) [10.1007/978-1-4615-4677-1]
\newblock R. G. Strongin and Y. D. Sergeyev,
\newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel
Algorithms},
\newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014)

\bibitem{zilinskTornGO}% (MR1100586) [10.1007/3-540-50871-6]
\newblock A. T\"orn and A. \v Zilinskas,
\newblock \emph{Global Optimization},
\newblock Springer, (1989)

\bibitem{Goryachih2017}
\newblock Goryachih, A.
\newblock \emph{A class of smooth modification of space-filling curves for global optimization
problems}
\newblock Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 57--65 (2017)

\bibitem{zhigljavskyRandGO}% (MR1187048) [10.1007/978-94-011-3436-1]
\newblock A. A. Zhigljavsky,
\newblock \emph{Theory of Global Random Search},
\newblock Kluwer Academic Publishers, Dordrecht (1991)

@@

\bibitem{globalizerSystem}
Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: \emph{A novel supercomputer software
system for solving time-consuming global optimization problems.} Numerical Algebra, Control
\& Optimization \textbf{8(1)}, 47--62 (2018)

@@

\bibitem{Jones}
Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M.
(eds.) \emph{The Encyclopedia of Optimization}, 2nd edn., pp. 725--735. Springer, Heidelberg
(2009)
